Large and Infinitary Quotient Inductive-Inductive Types

### 2020-01-10 draft/paper.tex

Algebraic notions of models of type theories are examples for this. Here,
initial models represent syntax, and initiality corresponds to induction on
syntax. Several variants have been used, from contextual categories \cite{gat}
and comprehension categories \cite{jacobs1993comprehension} to categories with
families \cite{Dybjer96internaltype}. Additionally, real, surreal \cite{HoTTbook}
and ordinal numbers \cite{lumsdaineShulman} can be defined as \emph{infinitary}
QIITs, which are represented as possibly infinitely branching inductive trees.

>>>>

Algebraic notions of models of type theories are examples for this. Here,
initial models represent syntax, and initiality corresponds to induction on
syntax. Several variants have been used, from contextual categories \cite{gat}
and comprehension categories \cite{jacobs1993comprehension} to categories with
families \cite{Dybjer96internaltype} (CwF).
%
Whereas the initial category can be given by first defining the type
of objects (the empty type) and then defining morphisms as a separate
inductive family, for the initial CwF one has to define $Ob$, $Mor$
and two extra sorts $Ty$ and $Tm$ \emph{inductive-inductively}, at the
same time. This is because there is a constructor of $Ob$ which refers
to $Ty$, constructors of $Ty$ refer to $Ob$ and $Sub$, and so on. This
is illustrated by the following excerpt of the signature for CwFs
extending the above signature for categories.
\begin{alignat*}{3}
  & Ty && : Ob\ra \U \\
  & empty  && : Ob \\
  & ext && : (I:Ob)\ra Ty\,I\ra Ob \\
  & \blank[\blank] && : Ty\,J\ra Mor\,I\,J\ra Ty\,I
\end{alignat*}

While signatures for type theory usually only contain finitary
constructors, important examples of QIITs have infinitary
constructors, such as the following signature for infinitaly branching
trees where the order of subtrees does not matter.
\begin{alignat*}{3}
  & Tree && : \U \\
  & leaf  && : Tree \\
  & node && : (\N\ra Tree)\ra Tree \\
  & perm && : (g :\N\ra Tree)(f : \N\ra\N)\ra \mathsf{isIso}\,f\ra \\
  & && \hspace{0.8em} node\,g = node (g\circ f)
\end{alignat*}
Here $node$ and $perm$ are infinitary constructurs, $perm$ expresses
that reordering subtrees by a permutation (a $\N\ra\N$ function which
is an isomorphism) does not matter.

Real-world examples examples of infinitary QIITs include the real, surreal numbers \cite{HoTTbook}, ordinal numbers \cite{lumsdaineShulman} and the partiality monad \cite{partiality}.
